\renewcommand{\abstractname}{Acknowledgements}
\begin{abstract}

First, I would like to thank my supervisor Dr. Ir. Ron van Lammeren for his help, advice and patience. He is very enthusiastic about my ideas and although he was often busy he found always time to help me. I also would like to thank Philip Wenting for provided the measuring devices and a measurement rollator. Take good care of our measurement rollator! 

The municipality of Amsterdam was very open and interested and I want to thank them for their time, advice and for providing me with the GBKA. Also the Waag Society was there to support me to go on and offered me a working place whenever I needed. 

More personally, I would like to thank Kristin Abraham, Andrei Mirt and Sabina Rosca for providing me with a sleeping place in Wageningen so I could work on my research. Also Andrei helped me getting started with \LaTeX  and solved problems whenever I needed. Thank you so much for your time! A very special thanks go to Kristin who proof read my whole thesis and gave a bunch of commentary which I really needed! Also thank you for the sweet cards and being there to support me, despite the distance. Keep on being great! Also I would like to thank my dad, Frans Boeijen, who did proof read my thesis while he actually did not have time for that at all. I am very grateful.  

Als laatse wil ik mijn opa, Jo Veld, bedanken voor de eeuwige inspiratie, moed en trots die je mij geeft.  

Niene Boeijen

March 2016
\end{abstract}

